---
title: "hw04"
author: "Maria Guo"
date: "October 10, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(gapminder)
library(tidyverse)
library(tidyr)
```

# General data reshaping and relationship to aggregation

## Make a tibble with one row per year and columns for life expectancy for two or more countries.

First, I will compare life expectancy between Canada, China and India by using filter and selection.

```{r}
lifeExp_Country <- gapminder %>%
        filter(country %in% c('Canada', 'India', 'China')) %>%
        select(country, year, lifeExp)

knitr::kable(lifeExp_Country, 
             align = 'c', format = 'markdown',
             col.names = c("Country", "Year", "Life Expectation"))
```

Then, to reshape these data, I will use spread(), which can make “wide” data longer.

```{r}
reshape_data <- lifeExp_Country %>%
        spread(country,lifeExp) 

knitr::kable(reshape_data,col.names = 
        c('Year','Canada', 'China', 'India'),
        align = 'c',
        format = 'markdown')
```

To make format better and easier to read. I will make a scatterplot.

```{r}
lifeExp_Country %>%
  ggplot(aes(x = year, y = lifeExp, colour=country)) + 
  geom_point() +
  scale_colour_discrete("Country") +
  scale_x_continuous(breaks = seq(1950, 2010, 10),
      labels = as.character(seq(1950, 2010, 10)),
      limits = c(1950, 2010),
      minor_breaks = NULL) +
  scale_y_continuous(breaks = seq(20, 85, 5),
      labels = as.character(seq(20, 85, 5)),
      limits = c(25, 85),
      minor_breaks = NULL) +
  theme_bw() + 
  labs(x = "Year", y = "Life Expectancy",
      title = "Comparison of Life Expectancies in Canada, India, and China")
```


## Compute some measure of life expectancy (mean? median? min? max?) for all possible combinations of continent and year. Reshape that to have one row per year and one variable for each continent. Or the other way around: one row per continent and one variable per year.

First, I will comoute the mean of life expectancy for each continent at each year.

```{r}
mean_lifeExp <- gapminder %>%
        group_by(continent, year) %>%
        mutate(lifeExp.mean = mean(lifeExp)) %>%
        select(continent, year, lifeExp.mean) %>%
        unique() %>%
        spread(continent,lifeExp.mean)

knitr::kable(mean_lifeExp,col.names = 
        c('Year','Africa', 'Americas','Europe','Oceania', 'Asia'),
        align = 'c',
        format = 'html', 
        caption = "Mean Life Expectancy")
```

To understand the data easier, I would compare mean life expectanc between 2 continents.


```{r}
mean_lifeExp <- gapminder %>%
        group_by(continent, year) %>%
        mutate(lifeExp.mean = mean(lifeExp)) %>%
        select(continent, year, lifeExp.mean) %>%
        unique() 

mean_lifeExp %>%
        filter(continent %in% c('Europe','Africa')) %>%
        ggplot(aes(x = year, y = lifeExp.mean, colour = continent)) +
        geom_point() +
        scale_x_continuous(breaks = seq(1950, 2010, 10),
        labels = as.character(seq(1950, 2010, 10)),
        limits = c(1950, 2010),
        minor_breaks = NULL) +
        scale_y_continuous(breaks = seq(20, 85, 5),
        labels = as.character(seq(20, 85, 5)),
        limits = c(25, 85),
        minor_breaks = NULL) +
        scale_colour_discrete(name = "continent") +
        theme_bw() +
        labs(x = "Year", y = "Mean Life Expectancy",
        title = "Comparison of Mean Life Expectancy by Year\nbetween Americas and Africa")
```



# Join, merge, look up

## Activity 1

First, I have to make a second data frame. This time, I use mean population in all continents by year. 
```{r}
second_data_frame <- gapminder %>% 
    filter(continent=="Asia") %>% 
    group_by(year,continent) %>% 
    summarize(mean_pop=mean(pop))

knitr::kable(second_data_frame, align = 'c',format = 'markdown', 
             col.names = c("Year", "Continent", "Mean Population"))
```

Then, I will play around with these two data frame by using various joint functions.

### Left Join
This will save all data in first data fram (gapminder), and add new column in second data frame to it. Therefore, at this time, continents, besides Asia, will have a N/A in new column.

```{r}
left_join(gapminder,second_data_frame, by = c("continent","year"))
```
```{r}
## Warning: Column `continent` joining factor and character vector, coercing
## into character vector
```


### Full Join

It retains all data in both data frame.
```{r}
full_join(gapminder,second_data_frame, by = c("continent","year"))
```
### Inner Join
It retains the data that only present in both data frame and add new columns to gapminder. So everything that associate with Asia is showed with data of mean population.
```{r}
inner_join(gapminder,second_data_frame, by = c("continent","year"))
```


### Semi Join
It is used to filter out the data that does not match with the one in second data frame, but does not add mean population to gapminder.

```{r}
semi_join(gapminder,second_data_frame, by = c("continent","year"))
```

### Anti Join

This is the inverse type of semi join. It only keep the data that is not present in second data frame. Therefore, all Asia data are removed and mean population is not added.
```{r}
anti_join(gapminder,second_data_frame, by = c("continent","year"))
```




